---
title: "Simple document"
output: github_document

    
---
```{r,echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(readr)
library(dplyr)
```

DS_problem1 
1.1read and clean the data
```{r}
Transit_data=read_csv("C:/Users/10145/Desktop/DS project/p8105_hw2_zx2425/p8105_hw2_zx2425/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c"))
names(Transit_data)
Transit_data=janitor::clean_names(Transit_data)
names(Transit_data)
tail(Transit_data)
data_c=select(Transit_data,line,station_name,station_latitude,station_longitude,route1:route11,entry,vending,entrance_type,ada)

```

DS_problem1 
1.2 Convert the entry variable's type
```{r}
data_c=mutate(data_c,entry=factor(entry,levels=c("YES","NO")))
```

DS_problem1 
1.3answer the questions
```{r}
distincted_station=distinct(data_c,station_name, .keep_all = TRUE)
filter_data=filter(distincted_station, ada==TRUE)
#356 rows
filter_data2=filter(distincted_station, vending=="NO")
filter_data2
```
1.3 the answer is:
The original data set which called Transit_data contains 32 variables. After standardizing the name of it, it including: division,line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entrance_type, entry, exit_only, vending, staffing, staff_hours, ada, ada_notes, free_crossover, north_south_street, east_west_street, corner, entrance_latitude, entrance_longitude, station_location, entrance_location.
The data set c is a cleaned dataset of original data of problem1. This data set contain 19 variables including line,station_name,station_latitude,station_longitude,route1:route11,entry,vending,entrance_type,ada.
I have standarded the name of the variables and select target variable to form a new dataset. it is 1868*19. I think those steps are data tidy.


```{r}
filter_data2 %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

```{r}
filter_data2 %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct


filter_data2 %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Explaination:
At last, we need to make a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).



Problem2

Read and clean the dataset of Mr_Trash_Wheel, The results as following:
```{r}
Mr_Trash_Wheel= 
    read_excel("C:/Users/10145/Desktop/DS project/p8105_hw2_zx2425/p8105_hw2_zx2425/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet = "Mr. Trash Wheel")%>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(
      sports_balls = as.integer(round(sports_balls))
      )
Mr_Trash_Wheel
```

Read and clean the dataset of Pro_Trash_Wheel, The results as following:
```{r}
Pro_Trash_Wheel = 
    read_excel("C:/Users/10145/Desktop/DS project/p8105_hw2_zx2425/p8105_hw2_zx2425/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet = "Professor Trash Wheel") %>% 
    janitor::clean_names() %>% 
    drop_na(dumpster) %>% 
    mutate(
      dumpster=as.double(dumpster)
    )
Pro_Trash_Wheel
```


```{r}
Mr_Trash_Wheel=Mr_Trash_Wheel[ , -which(colnames(Mr_Trash_Wheel) %in% c("x15","x16"))]
Mr_Trash_Wheel= mutate(Mr_Trash_Wheel,name=as.character("Mr. Trash Wheel"))
Pro_Trash_Wheel=mutate(Pro_Trash_Wheel,name=as.character("Professor Trash Wheel"))
Mr_Trash_Wheel= mutate(Mr_Trash_Wheel,year=as.double(year))
Pro_Trash_Wheel= mutate(Pro_Trash_Wheel,year=as.double(year))
TOTAL= bind_rows(Mr_Trash_Wheel,Pro_Trash_Wheel)
```


```{r}
sum(Pro_Trash_Wheel$weight_tons)
filter_Mr_Trash_Wheel=dplyr::filter(Mr_Trash_Wheel,year==2020)
sum(filter_Mr_Trash_Wheel$sports_balls)
```

Explaination:
The total weight of trash collected by Professor Trash Wheel is ```r sum(Pro_Trash_Wheel$weight_tons) ```
The total number of sports balls collected by Mr. Trash Wheel in 2020 is ```r
`sum(filter_Mr_Trash_Wheel$sports_balls)```

Problem 3

Problem3
Read three datasets:
```{r}
unemployment=read_csv("C:/Users/10145/Desktop/DS project/p8105_hw2_zx2425/p8105_hw2_zx2425/fivethirtyeight_datasets/unemployment.csv")
pol=read_csv("C:/Users/10145/Desktop/DS project/p8105_hw2_zx2425/p8105_hw2_zx2425/fivethirtyeight_datasets/pols-month.csv")
snp=read_csv("C:/Users/10145/Desktop/DS project/p8105_hw2_zx2425/p8105_hw2_zx2425/fivethirtyeight_datasets/snp.csv")
```

manipulate the pols_moth dataset
```{r} 
pol = separate(pol,mon, into= c("year","month",'day'),sep= "-")
pol=mutate(pol,
  month = recode(month, "01" = "January", "02" = "February", "03"="March", "04"="April","05"="May","06"="June", "07"="July", "08"="August","09"="September","10"="October","11"="November", "12"="December")
)
pol=mutate(pol,president=ifelse(prez_gop==1, "prez_gop",ifelse(prez_gop==0,"prez_dem","other") )
           )
```

manipulate the pols_moth dataset
```{r}
pol=select(pol, -day)
pol=select(pol, -prez_gop)
pol=select(pol, -prez_dem)
```


manipulate the snp dataset
```{r}
snp = separate(snp,date, into= c("month","day",'year'),sep= "/")
snp=mutate(snp, month = recode(month, "1" = "January", "2" = "February", "3"="March", "4"="April","5"="May","6"="June", "7"="July", "8"="August","9"="September","10"="October","11"="November", "12"="December")
)
snp=select(snp, -day)
snp=mutate(snp,year=as.integer(year))
```

translate the formate of the two datasets in order to combine them together:
```{r}
pol=mutate(pol,year=as.integer(year))
data_d=pivot_longer(unemployment,Jan:Dec,names_to="month",values_to="amount")
data_d= data_d %>%
  janitor::clean_names()
data_d=mutate(data_d,year=as.integer(year))
snp=mutate(snp, year = ifelse(year<20,year + 2000,year + 1900))
snp=arrange(snp,year,month)
snp=relocate(snp, year, month,close)
```

jion them together
```{r}
library(dplyr)
data_e=left_join(pol,snp,by=c("year","month"))
data_f=left_join(data_e,data_d,by=c("year","month"))

```

the content of the initial pols-month dataframe contains 9 variables, including:mon      prez_gop gov_gop sen_gop rep_gop prez_dem gov_dem sen_dem rep_dem.822 multiply 9. After processing, they become to year month gov_gop sen_gop rep_gop gov_dem sen_dem rep_dem president.822*9 The range of the year of pols_moth dataframe is 1947-2015.
the content of the initial snp dataframe contains 2 variables, including:data close. 787*2 After processing, they become to month year close.787 multiply 3. The range of the year of snp dataframe is 1950-2015.

the content of the combined dataframe：data_f contains 11 variables, including:year month  gov_gop sen_gop rep_gop gov_dem sen_dem rep_dem president close amount. 822*11. The range of the combined dataframe data_f is 1947-2015.



